---
title: "Statistical Analysis"
description: |
  A sample report using PCA.
date: 2021-02-28
output:
  distill::distill_article:
    code_folding: true
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(here)
library(ggfortify)

```

### 1. Introduction
This mini-report uses data compiled by Zander Venter using Google Earth Engine to summarize environmental and climatic variables for countries across the globe. The report explores the relationships among a subset of those variables, including country elevation, cropland cover, tree canopy cover, average annual rainfall and temperature, wind and cloudiness, using principal components analysis. The results are visualized in a biplot and interpreted in a brief summary.

### 2. Exploratory Findings & Summary

```{r, results=FALSE, fig.cap="Relationships among select environmental and climatic variables for countries across the globe. Vectors indicate loadings of variables along each principal component. Data: Venter, 2018"}
# read in data
world_env_vars <- read.csv(here("data", "world_env_vars.csv"))

# wrangling + pca 
world_env_vars_pca <- world_env_vars %>% 
  select(elevation, cropland_cover, tree_canopy_cover, rain_mean_annual, temp_mean_annual, wind, cloudiness) %>% # retains only the variables elevation, cropland and tree canopy cover, average annual rain and temperature, wind, & cloudiness
  drop_na() %>% # removes na values for the above variables
  scale() %>% # scales data to create equal units of measurement
  prcomp() # principal components analysis

# to see loadings
world_env_vars_pca$rotation

# biplot
world_env_vars_complete <- world_env_vars %>% 
  drop_na(elevation, cropland_cover, tree_canopy_cover, rain_mean_annual, temp_mean_annual, wind, cloudiness) # creates data subset to match that used for pca

autoplot(world_env_vars_pca,
         data = world_env_vars_complete,
         colour = 'seagreen2',
         loadings = TRUE,
         loadings.label = TRUE,
         loadings.colour = "black",
         loadings.label.colour = "black",
         loadings.label.vjust = -.5,
         loadings.label.hjust = 1,
         loadings.label.size = 3.25) +
  theme_minimal() +
  labs(x = "Component 1 (40.33%)", y = "Component 2 (20.95%)")
```

#### Summary
Major takeaways from the PCA include:

- Tree canopy cover and average annual rainfall are positively correlated.
- Cropland cover and elevation are positively correlated.
- Cropland cover and average annual temperature are negatively correlated.
- Wind and tree canopy cover are negatively correlated.
- Cloudiness and cropland cover are minimally correlated.
- There are no observable clusters of observations and only a couple of outliers.
- Cloudiness, tree canopy cover, and average annual rainfall are heavily loaded in the direction of the first principal component.
- Average annual temperature and cropland cover are heavily loaded in the direction of the second principal component.
- Over 60% of total variance is represented.
